Curvature approximation on surfaces -- A Discrete Exterior Calculus Approach

### 2014-12-08 article/main.tex
...

        A more ``FEM-like'' manner is to decompose \eqref{eqDisLaplace} in parts, which can be computed on all (triangle) elements without
        knowledge of their neighbours.
        For this we formulate the operator on the Voronoi cells to cancel the scale factor \( \left| \star v \right|^{-1} \) 
        in \eqref{eqDisLaplace}, split the sum and use a local vertex indexing on all elements:
        \begin{align} \label{eqVoronoiLaplace}
                \left\langle *\Delta_{B} f , \star v_{i} \right\rangle
                               &= \sum_{\sigma^{1}=\left[ v_{i}, v_{j} \right]} 
                     \frac{\left| \star\sigma^{1} \right|}{\left| \sigma^{1} \right|}
                      \left( f(v_{j}) - f(v_{i}) \right)\\
                               &= \sum_{\substack{\sigma^{2}=\left[ v^{\sigma^{2}}_{0},v^{\sigma^{2}}_{1},v^{\sigma^{2}}_{2} \right]
                              \\
                                               v^{\sigma^{2}}_{0} = v_{i}}}
                  \sum_{l=1,2} C^{\sigma^{2}}_{0,l} 
                          \left( f^{\sigma^{2}}_{l} -  f^{\sigma^{2}}_{0}\right) \formpunkt \label{eqLokalLaplace}
        \end{align}
        \( f^{\sigma^{2}}_{l}\in\R \) are the evaluations of \( f \) on the local vertices \( v^{\sigma^{2}}_{l} \) of the element \(
        \sigma^{2}\).
        Figure \ref{figLocalIndexing} clarify this setting.
        The coefficients \( C^{\sigma^{2}}_{0,l}\in\R \) are generally defined by
        \begin{align}
               C^{\sigma^{2}}_{k,l} &:= C^{\sigma^{2}}_{l,k}
                            := \frac{\left| \star \left[ v^{\sigma^{2}}_{k}, v^{\sigma^{2}}_{l} \right] 
                                            \cap \sigma^{2}\right|}
                                   {\left| \left[ v^{\sigma^{2}}_{k}, v^{\sigma^{2}}_{l} \right] \right|}\formpunkt
        \end{align}
        \begin{figure}
          \centering
          \includegraphics[width=.45\textwidth]{bilder/tikz/elementSummeKante.eps}
          \caption{Two adjacent (triangle) elements share a common edge \( \sigma^{1}=\left[ v_{i}, v_{j} \right] \).
                   The vertices are local indexed on these elements.}
          \label{figLocalIndexing}
        \end{figure}
        Hence we can extract from \eqref{eqLokalLaplace} a local coefficients matrix
        \begin{align}
          A^{\sigma^{2}} 
              &:= {\footnotesize\begin{bmatrix}
                    -\left( C^{\sigma^{2}}_{01} + C^{\sigma^{2}}_{02}\right) & C^{\sigma^{2}}_{01} & C^{\sigma^{2}}_{02} \\
                    C^{\sigma^{2}}_{01} & -\left( C^{\sigma^{2}}_{01} + C^{\sigma^{2}}_{12}\right) & C^{\sigma^{2}}_{12} \\
                    C^{\sigma^{2}}_{02} & C^{\sigma^{2}}_{12} & -\left( C^{\sigma^{2}}_{02} + C^{\sigma^{2}}_{12}\right)
                  \end{bmatrix}} 
        \end{align}
        which holds all computation parts for the local vertices.
        All \( A^{\sigma^{2}} \) can be assembly in a familiar way to get a sparse global matrix 
        \( A\in\R^{N_{\sigma^{0}} \times N_{\sigma^{0}}} \), where \( N_{\sigma^{0}} \) are the number of vertices in \( K \).
        At the end, for \( f_{h}\in\R^{N_{\sigma^{0}}} \), the vector of evaluations for \( f \) 
        (resp. vector of DOFs\footnote{Degree Of Freedom}), 
        \( Af_{h} \) is the result of \eqref{eqVoronoiLaplace} on all global vertices.

@@

\section{Curvature approximations}
  \subsection{Curvature vector} \label{secCurvatureVector}
  \subsection{Weingarten map} \label{secWeingarten}

@@

>>>>

        A more ``FEM-like'' manner is to decompose \eqref{eqDisLaplace} in parts, which can be computed on all (triangle) elements without
        knowledge of their neighbours.
        For this we formulate the operator on the Voronoi cells to cancel the scale factor \( \left| \star v \right|^{-1} \) 
        in \eqref{eqDisLaplace}, split the sum and use a local vertex indexing on all elements:
        \begin{align} \label{eqVoronoiLaplace}
                \left\langle *\Delta_{B} f , \star v_{i} \right\rangle
                               &= \sum_{\sigma^{1}=\left[ v_{i}, v_{j} \right]} 
                     \frac{\left| \star\sigma^{1} \right|}{\left| \sigma^{1} \right|}
                      \left( f(v_{j}) - f(v_{i}) \right)\\
                               &= \sum_{\substack{\sigma^{2}=\left[ v^{\sigma^{2}}_{0},v^{\sigma^{2}}_{1},v^{\sigma^{2}}_{2} \right]
                              \\
                                               v^{\sigma^{2}}_{0} = v_{i}}}
                  \sum_{l=1,2} C^{\sigma^{2}}_{0,l} 
                          \left( f^{\sigma^{2}}_{l} -  f^{\sigma^{2}}_{0}\right) \formpunkt \label{eqLokalLaplace}
        \end{align}
        \( f^{\sigma^{2}}_{l}\in\R \) are the evaluations of \( f \) on the local vertices \( v^{\sigma^{2}}_{l} \) of the element \(
        \sigma^{2}\).
        Figure \ref{figLocalIndexing} clarify this setting.
        The coefficients \( C^{\sigma^{2}}_{0,l}\in\R \) are generally defined by
        \begin{align} \label{eqLokalMatrixLaplace}
               C^{\sigma^{2}}_{k,l} &:= C^{\sigma^{2}}_{l,k}
                            := \frac{\left| \star \left[ v^{\sigma^{2}}_{k}, v^{\sigma^{2}}_{l} \right] 
                                            \cap \sigma^{2}\right|}
                                   {\left| \left[ v^{\sigma^{2}}_{k}, v^{\sigma^{2}}_{l} \right] \right|}\formpunkt
        \end{align}
        \begin{figure}
          \centering
          \includegraphics[width=.45\textwidth]{bilder/tikz/elementSummeKante.eps}
          \caption{Two adjacent (triangle) elements share a common edge \( \sigma^{1}=\left[ v_{i}, v_{j} \right] \).
                   The vertices are local indexed on these elements.}
          \label{figLocalIndexing}
        \end{figure}
        Hence we can extract from \eqref{eqLokalLaplace} a local coefficients matrix
        \begin{align}
          A^{\sigma^{2}} 
              &:= {\footnotesize\begin{bmatrix}
                    -\left( C^{\sigma^{2}}_{01} + C^{\sigma^{2}}_{02}\right) & C^{\sigma^{2}}_{01} & C^{\sigma^{2}}_{02} \\
                    C^{\sigma^{2}}_{01} & -\left( C^{\sigma^{2}}_{01} + C^{\sigma^{2}}_{12}\right) & C^{\sigma^{2}}_{12} \\
                    C^{\sigma^{2}}_{02} & C^{\sigma^{2}}_{12} & -\left( C^{\sigma^{2}}_{02} + C^{\sigma^{2}}_{12}\right)
                  \end{bmatrix}} 
        \end{align}
        which holds all computation parts for the local vertices.
        All \( A^{\sigma^{2}} \) can be assembly in a familiar way to get a sparse global matrix 
        \( A\in\R^{N_{\sigma^{0}} \times N_{\sigma^{0}}} \), where \( N_{\sigma^{0}} \) are the number of vertices in \( K \).
        At the end, for \( f_{h}\in\R^{N_{\sigma^{0}}} \), the vector of evaluations for \( f \) 
        (resp. vector of DOFs\footnote{Degree Of Freedom}), 
        \( Af_{h} \) is the result of \eqref{eqVoronoiLaplace} on all global vertices.

@@

\section{Curvature approximations}
  \subsection{Curvature vector} \label{secCurvatureVector}
    The curvature vector \( \vec{H} \) is given by
    \begin{align}
      \vec{H} = 2 H \vec{\nu}
    \end{align}
    for the mean curvature \( H \) and normals \( \vec{\nu} \). 
    In this paper, the mean curvature is set as the arithmetic mean of the two principal curvatures and
    \( H \) should be all over positive for a sphere (i.e. \( H\equiv1 \) on \( \mathds{S}^{2} \)).
    \( H \) and \( \vec{H} \) are extrinsic curvature values.
    This means, that an two dimensional ``inhabitant'' of the surface can't determine this kind of curvature 
    in contrast to the Gaussian curvatures.
    With the inclusion map \( \vec{x}:= \R^{3}\supset M \hookrightarrow \R^{3} \), which is the position vector of \( M \) in \( \R^{3} \),
    the curvature vector can be calculated by (see \cite[Ch. 5]{chen}, \cite[Ch. 4.5]{flanders} or \cite{elliott})
    \begin{align} \label{eqCurvatureVector}
      -\Delta_{B}\vec{x} = \vec{H} \formpunkt
    \end{align}
    (Note that \( \vec{x} \) is an isometric immersion with the common induced metric from \( \R^{3} \).)
    The equation \eqref{eqCurvatureVector} is component-wise to read and can be discretized with 
    discrete Laplace-Beltrami operator in \eqref{eqDisLaplace}, 
    where \( f \) becomes the \mh{i}{th} coordinate functions \( x^{i} \) for
    \( i=0,1,2 \).
    Hence, we can easily build the local coefficients matrices for the elements \( \sigma^{2} \) 
    with \( A^{\sigma^{2}} \) from \eqref{eqLokalMatrixLaplace}:
    \begin{align}
      \begin{bmatrix}
        A^{\sigma^{2}} & 0 & 0 \\
        0 & A^{\sigma^{2}} & 0 \\
        0 & 0 & A^{\sigma^{2}}
      \end{bmatrix}
    \end{align}
    to determine the left-hand side of the system
    \begin{align} \label{eqDisCurvatureVector}
      \left\langle *\Delta_{B} x^{i}, \star v \right\rangle
        &= \left\langle *H^{i}, \star v \right\rangle \formpunkt
    \end{align}
    for all vertices \( v\in K \) and \( i=0,1,2 \).
    \( H^{i}\in \R \) are the components of the curvature vector.
    The right-hand side of \eqref{eqDisCurvatureVector} can be locally represented by the decomposed Voronoi area 
    \begin{align}
      C^{\sigma^{2}}_{i} = \left|  \right|
    \end{align}
    around \( v \).

  \subsection{Weingarten map} \label{secWeingarten}

@@

  \begin{figure}
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/K.jpg}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/H.jpg}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/L2K.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/L2H.eps}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/LMaxK.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/torus/LMaxH.eps}
    \end{minipage}
    \caption{Torus: Top row: Gaussian curvature \( K \) (left) and mean curvature \( H \) (right).
                              Middle and bottom row: Relative \( L_{2} \) resp. \( L_{\infty} \) error for \( K \) (left) and
                                                     \( H \) (right) for different \( h \) in a log-log-plot.}
    \label{figTorus}
  \end{figure}
